---
title: "20120924_exam_rmed901_sindre_grindheim"
author: "Sindre Grindheim"
date: "9/25/2021"
output: html_document
---

First I install and load data packages needed. 
```{r}
install.packages("tidyverse")
install.packages("here")
install.packages("tibble")
```
```{r}
library(tidyverse)
library(here)
library(tibble)
```


Import exam data file
```{r}
covid <-
  read.delim("/Users/sindregrindheim/Desktop/RMED901/2021_09_24_RMED901_Exam_Sindre_Grindheim/DATA/exam_data.txt", sep = "\t")
print(covid)
```

```{r}
summarise(covid)
head(covid)
tail(covid)
glimpse(covid)
count(covid)
```
Create tibble
```{r}
covid <- as_tibble(covid)
```



1st column includes both name and number, possible ID number. These have to be separated. Use separate function, naming the column needed to be split, create a new vector with the new column names, and name the seperate marker. 

```{r}
covid <-
  covid %>% 
  separate(subject, c("id_number", "first_name", "surname"), "-")
```

Make column names uniform related to denominator. Use underscore for all

```{r}
covid <-
  rename(covid, pan_day = pan.day)
```

Create separate column for gender. Mutate to create a new column based on missing value in the age_female group. This one did not work, not sure why. 
```{r}
mutate(covid, gender = if_else(age_female == NA, "F", "F", "M"))
```
Created seperate column for gender using values >= 0 instead of basing it on NA. Worked, however, less elegant.

```{r}
covid <-
  covid %>% 
  mutate(gender = case_when(
    age_female >= 0 ~ "F", age_male >= 0 ~ "M")
  )
```


```{r}
glimpse(covid)
summarise(covid)
```


```{r}
install.packages(c("finalfit", "skimr", "naniar"))

```

Checking for missing values. ff_glimpse gives a percentage. a max on col_rec_tat of 61370.2 where the mean is 7.2 indicates an outlier/plotting mistake
```{r}
is.na(covid)
ff_glimpse(covid)
skimr::skim(covid)
```

find outliers. Making histiograms of the continous variables.

```{r}
library(ggplot2)

hist_age_m <-
ggplot(covid) +
  aes(x = age_male) +
  geom_histogram(bins = 30L, fill = "#0c4c8a") +
  theme_minimal()
```


```{r}
hist_age_f <-
ggplot(covid) +
  aes(x = age_female) +
  geom_histogram(bins = 30L, fill = "#0c4c8a") +
  theme_minimal()
```
```{r}
hist_pan_day <-
ggplot(covid) +
  aes(x = pan_day) +
  geom_histogram(bins = 30L, fill = "#0c4c8a") +
  theme_minimal()
```


```{r}
hist_rec_ver_tat <-
ggplot(covid) +
  aes(x = rec_ver_tat) +
  geom_histogram(bins = 30L, fill = "#0c4c8a") +
  theme_minimal()
```


```{r}
hist_col_rec_tat <-
ggplot(covid) +
  aes(x = col_rec_tat) +
  geom_histogram(bins = 30L, fill = "#0c4c8a") +
  theme_minimal()
```


```{r}
hist_ct_result <-
ggplot(covid) +
  aes(x = ct_result) +
  geom_histogram(bins = 30L, fill = "#0c4c8a") +
  theme_minimal()
```


time spent waiting for the test result in the start vs the end of the pandemic. Development. 

```{r}
covid %>% 
  ggplot(aes(x = pan_day, y = col_rec_tat)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ggpubr::stat_cor()
  


```
remove outlier >50. Result did not change regardles of level set for removing outlier (>24)
Conclusion: Time from test to result did not change from the start to the end

```{r}
filter(covid, col_rec_tat < 50) %>%
   ggplot(aes(x = pan_day, y = col_rec_tat)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ggpubr::stat_cor()
  
```
Number of tests by gender. More tests among females

```{r}
covid %>% 
count(gender)
```

lin reg of test results time stratified by gender


```{r}
  filter(covid, col_rec_tat < 50) %>%
   ggplot(aes(x = pan_day, y = col_rec_tat)) +
    geom_point(aes(color = gender)) +
  geom_smooth(aes(linetype = gender), method = "lm") +
  facet_grid(gender ~ .)

  
```



Are there more positive tests in the drive-through?
Filter out the positive results, count the freq of drive through or not among these. 
Result: more positive test in drivethrough

```{r}
  pos <-
    filter(covid, result == "positive")

```

```{r}
pos %>% 
  count(drive_thru_ind)
```

Is the age of the individual associated with whether the test result is negative or positive?

```{r}
 filter(covid, col_rec_tat < 50) %>%
   ggplot(aes(y = age_female, x = result)) +
    geom_point(aes(color = result)) +
  geom_jitter(alpha = 0.4)
  
```

```{r}
covid %>% 
  ggplot(aes(x = result, y = age_female)) +
  geom_boxplot() +
  ggpubr::stat_compare_means()
```

```{r}
covid %>% 
  ggplot(aes(x = result, y = age_male)) +
  geom_boxplot() +
  ggpubr::stat_compare_means()
```
converting into logarithm

```{r}
covid %>% 
  mutate(log_age_f = log(age_female)) %>% 
  ggplot(aes(x = log_age_f)) +
  geom_histogram() +
  facet_grid(~result)
```


running anova
```{r}
covid %>% 
  mutate(agef = log(age_female)) %>% 
  aov(agef ~ result, data = .)

```